---
title: 'SoundJudgement: frequentist and Bayesian Analysis'
author: "Filippo Spinelli"
date: "2026-09-19"
output: html_document
---

# ============================================================
# Dataset: ABtestdata  
# Variables:
#   pairwise              : pair ID 1–766
#   happyBright           : 1 = Happy & Bright music, 0 = Dark & Ominous
#   hit                   : 1 = believes children will be hit, 0 = not
#   effectiveRailwayVideo : 1–10 rating of railway safety video
# ============================================================

# ---- PACKAGES -----------------------------------------------

```{r}
required <- c("tidyverse", "ggplot2", "patchwork", "scales", "readxl")
to_install <- required[!required %in% rownames(installed.packages())]
if (length(to_install)) install.packages(to_install)
lapply(required, library, character.only = TRUE)
```

# ============================================================
# SECTION 1: DESCRIPTIVE ANALYSIS & ERROR DETECTION
# ============================================================

```{r}
library(readxl)
ABtestdata <- read_excel("C:/Users/User1/OneDrive - Alma Mater Studiorum Università di Bologna/Business Analytics/Business statistics/Bayesian project/ABtestdata.xlsx")

cat("=== 1. DATASET OVERVIEW ===\n")
glimpse(ABtestdata)

cat("\n=== 1.1 SUMMARY STATISTICS (RAW) ===\n")
print(summary(ABtestdata))
```

# ---- Identify errors ----------------------------------------
# Expected ranges: hit in {0,1}, effectiveRailwayVideo in {1,...,10}

```{r}
error_rows <- ABtestdata %>%
  mutate(row_id = row_number()) %>%
  filter(!hit %in% c(0, 1) |
         effectiveRailwayVideo < 1 |
         effectiveRailwayVideo > 10)

cat("\n=== 1.2 ERROR ROWS DETECTED ===\n")
print(error_rows)
```

# Expected output:
#   row 1522: pairwise=761, happyBright=1, hit=-1,  effectiveRailwayVideo=9
#   row 1524: pairwise=762, happyBright=1, hit=0,   effectiveRailwayVideo=-10

# ERROR 1 — hit = -1: impossible value; true value unknown → REMOVE
# ERROR 2 — effectiveRailwayVideo = -10: likely accidental minus sign;
#            but since we cannot be certain it was 10, we → REMOVE
# Both rows represent < 0.14% of the data so removal has negligible impact.

```{r}
ABclean <- ABtestdata %>%
  filter(hit %in% c(0, 1),
         effectiveRailwayVideo >= 1,
         effectiveRailwayVideo <= 10)

cat(sprintf("\nOriginal: %d rows  |  After cleaning: %d rows  |  Removed: %d\n",
            nrow(ABtestdata), nrow(ABclean), nrow(ABtestdata) - nrow(ABclean)))
```

# ---- Descriptive stats by group -----------------------------

```{r}
ABclean <- ABclean %>%
  mutate(Music = ifelse(happyBright == 1, "Happy & Bright", "Dark & Ominous"))

desc_stats <- ABclean %>%
  group_by(Music) %>%
  summarise(
    n               = n(),
    n_hit           = sum(hit),
    prop_hit        = mean(hit),
    mean_railway    = mean(effectiveRailwayVideo),
    sd_railway      = sd(effectiveRailwayVideo),
    median_railway  = median(effectiveRailwayVideo),
    .groups = "drop"
  )

cat("\n=== 1.3 DESCRIPTIVE STATISTICS BY GROUP ===\n")
print(desc_stats)
```

# ── Binary variables: frequency tables ──────────────────────

```{r}
cat("=== happyBright ===\n")
table(ABclean$happyBright) %>% 
  prop.table() %>% 
  cbind(count = table(ABclean$happyBright), proportion = .) %>% 
  print()

cat("\n=== hit ===\n")
table(ABclean$hit) %>% 
  prop.table() %>% 
  cbind(count = table(ABclean$hit), proportion = .) %>% 
  print()
```

# ── Continuous variable: proper numeric summary ──────────────

```{r}
cat("\n=== effectiveRailwayVideo ===\n")
ABclean %>% 
  summarise(
    n       = n(),
    mean    = mean(effectiveRailwayVideo),
    sd      = sd(effectiveRailwayVideo),
    min     = min(effectiveRailwayVideo),
    Q1      = quantile(effectiveRailwayVideo, 0.25),
    median  = median(effectiveRailwayVideo),
    Q3      = quantile(effectiveRailwayVideo, 0.75),
    max     = max(effectiveRailwayVideo)
  ) %>% print()
```

# ---- VISUALISATIONS -----------------------------------------

# -- PLOT 1: Flagging the two error rows ----------------------

```{r}
p_errors <- ABtestdata %>%
  mutate(
    row_id    = row_number(),
    has_error = !hit %in% c(0, 1) |
                effectiveRailwayVideo < 1 |
                effectiveRailwayVideo > 10
  ) %>%
  ggplot(aes(x = row_id, y = effectiveRailwayVideo, colour = has_error)) +
  geom_point(size = 0.5, alpha = 0.4) +
  geom_point(data = . %>% filter(has_error),
             size = 5, shape = 8, colour = "red") +
  annotate("text",
           x = c(1522, 1524),
           y = c(-10, -10),
           label = c("hit = -1\n(pair 761)", "rating = -10\n(pair 762)"),
           colour = "red", size = 3.2, vjust = 1.2, hjust = 0.5) +
  scale_colour_manual(values = c("FALSE" = "#95a5a6", "TRUE" = "red"),
                      labels = c("Valid", "Error"),
                      guide  = guide_legend(override.aes = list(size = 3))) +
  labs(title    = "Data Validation: Railway Video Ratings",
       subtitle = "Red stars mark the two erroneous observations",
       x = "Row index", y = "effectiveRailwayVideo",
       colour = "Status") +
  theme_minimal(base_size = 13)

print(p_errors)
```

# -- PLOT 2: Proportion of 'hit' by music group ---------------

```{r}
p_hit_bar <- ABclean %>%
  count(Music, hit) %>%
  group_by(Music) %>%
  mutate(prop = n / sum(n),
         label_pct = scales::percent(prop, accuracy = 0.1),
         Hit = ifelse(hit == 1, "Will be hit", "Won't be hit")) %>%
  ggplot(aes(x = Music, y = prop, fill = Hit)) +
  geom_col(width = 0.55, colour = "white", linewidth = 0.4) +
  geom_text(aes(label = label_pct),
            position = position_stack(vjust = 0.5),
            size = 4.5, colour = "white", fontface = "bold") +
  scale_fill_manual(values = c("Will be hit"  = "#e74c3c",
                               "Won't be hit" = "#2ecc71")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title    = "Q1: Proportion believing children will be hit",
       subtitle = "By music condition (cleaned data)",
       x = NULL, y = "Proportion", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

print(p_hit_bar)
```

# -- PLOT 3: Railway rating distributions ---------------------

```{r}
p_rail_hist <- ABclean %>%
  ggplot(aes(x = effectiveRailwayVideo, fill = Music)) +
  geom_histogram(binwidth = 1, position = "dodge",
                 colour = "white", alpha = 0.85) +
  geom_vline(data = desc_stats,
             aes(xintercept = mean_railway, colour = Music),
             linetype = "dashed", linewidth = 1.1) +
  scale_fill_manual(values  = c("Happy & Bright" = "#f39c12",
                                "Dark & Ominous" = "#2c3e50")) +
  scale_colour_manual(values = c("Happy & Bright" = "#f39c12",
                                 "Dark & Ominous" = "#2c3e50")) +
  scale_x_continuous(breaks = 1:10) +
  labs(title    = "Q2: Railway Video Effectiveness Ratings",
       subtitle = "Dashed lines = group means",
       x = "Rating (1–10)", y = "Count",
       fill = "Music", colour = "Music") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

p_rail_violin <- ABclean %>%
  ggplot(aes(x = Music, y = effectiveRailwayVideo, fill = Music)) +
  geom_violin(alpha = 0.45, width = 0.8, colour = NA) +
  geom_boxplot(width = 0.22, alpha = 0.9,
               outlier.shape = 21, outlier.size = 1.5) +
  scale_fill_manual(values = c("Happy & Bright" = "#f39c12",
                               "Dark & Ominous" = "#2c3e50")) +
  scale_y_continuous(breaks = 1:10) +
  labs(title = "Q2: Rating Distribution (violin + boxplot)",
       x = NULL, y = "Rating (1–10)", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")

print(p_rail_hist + p_rail_violin)
```

# ============================================================
# SECTION 2: Q1 — DOES MUSIC AFFECT P(CHILDREN WILL BE HIT)?
#   Binary outcome → Beta-Binomial Bayesian + proportion test
# ============================================================

```{r}
HB <- ABclean %>% filter(happyBright == 1) %>% pull(hit)
DO <- ABclean %>% filter(happyBright == 0) %>% pull(hit)

n_HB <- length(HB); k_HB <- sum(HB)
n_DO <- length(DO); k_DO <- sum(DO)

cat(sprintf("\n=== 2. Q1 — HIT PROPORTIONS ===\n"))
cat(sprintf("Happy & Bright : %d/%d  = %.4f\n", k_HB, n_HB, k_HB / n_HB))
cat(sprintf("Dark & Ominous : %d/%d  = %.4f\n", k_DO, n_DO, k_DO / n_DO))
```


# ---- 2.1 Frequentist: two-proportion z-test -----------------

```{r}
freq_q1 <- prop.test(c(k_HB, k_DO), c(n_HB, n_DO),
                     alternative = "two.sided", correct = FALSE)
cat("\n--- Frequentist Two-Proportion Test ---\n")
print(freq_q1)
```

# ---- 2.2 Bayesian: Beta-Binomial ----------------------------
# Non-informative prior: Beta(1, 1) = Uniform[0,1] for both groups

```{r}
prior_a <- 1; prior_b <- 1

# Posterior update: Beta(alpha + successes, beta + failures)
a_HB_post <- prior_a + k_HB;   b_HB_post <- prior_b + (n_HB - k_HB)
a_DO_post <- prior_a + k_DO;   b_DO_post <- prior_b + (n_DO - k_DO)

cat(sprintf("\n--- Bayesian Beta-Binomial (full data) ---\n"))
cat(sprintf("Posterior HB : Beta(%d, %d)  =>  mean = %.4f\n",
            a_HB_post, b_HB_post, a_HB_post / (a_HB_post + b_HB_post)))
cat(sprintf("Posterior DO : Beta(%d, %d)  =>  mean = %.4f\n",
            a_DO_post, b_DO_post, a_DO_post / (a_DO_post + b_DO_post)))
```

# Monte Carlo estimate of P(p_HB > p_DO)

```{r}
set.seed(42)
N_MC <- 300000
p_HB_mc <- rbeta(N_MC, a_HB_post, b_HB_post)
p_DO_mc  <- rbeta(N_MC, a_DO_post, b_DO_post)
prob_HB_wins_q1 <- mean(p_HB_mc > p_DO_mc)

cat(sprintf("P(p_HB > p_DO | data)   = %.4f\n", prob_HB_wins_q1))
cat(sprintf("95%% Credible Interval HB : [%.4f, %.4f]\n",
            qbeta(0.025, a_HB_post, b_HB_post), qbeta(0.975, a_HB_post, b_HB_post)))
cat(sprintf("95%% Credible Interval DO : [%.4f, %.4f]\n",
            qbeta(0.025, a_DO_post, b_DO_post), qbeta(0.975, a_DO_post, b_DO_post)))
```

# ---- VIZ: Posterior densities Q1 ----------------------------

```{r}
theta_grid <- seq(0.1, 0.7, length.out = 2000)

post_df_q1 <- data.frame(
  theta = rep(theta_grid, 3),
  density = c(
    dbeta(theta_grid, prior_a,   prior_b),
    dbeta(theta_grid, a_HB_post, b_HB_post),
    dbeta(theta_grid, a_DO_post, b_DO_post)
  ),
  group = rep(c("Prior  Beta(1, 1)",
                "Posterior: Happy & Bright",
                "Posterior: Dark & Ominous"), each = 2000)
)

p_posterior_q1 <- ggplot(post_df_q1,
                         aes(x = theta, y = density,
                             colour = group, linetype = group)) +
  geom_line(linewidth = 1.2) +
  scale_colour_manual(values = c("Prior  Beta(1, 1)"          = "#95a5a6",
                                 "Posterior: Happy & Bright"  = "#f39c12",
                                 "Posterior: Dark & Ominous"  = "#2c3e50")) +
  scale_linetype_manual(values = c("Prior  Beta(1, 1)"         = "dashed",
                                   "Posterior: Happy & Bright" = "solid",
                                   "Posterior: Dark & Ominous" = "solid")) +
  annotate("label",
           x = 0.62, y = max(dbeta(theta_grid, a_HB_post, b_HB_post)) * 0.75,
           label = sprintf("P(p_HB > p_DO) = %.3f", prob_HB_wins_q1),
           size = 4.5, colour = "#c0392b", fill = "white", label.size = 0.4,
           fontface = "bold") +
  labs(title    = "Q1: Posterior distributions — P(children will be hit)",
       subtitle = "Beta-Binomial model with non-informative prior Beta(1, 1)",
       x = "Probability of 'hit'", y = "Posterior density",
       colour = "Distribution", linetype = "Distribution") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        legend.text = element_text(size = 10))

print(p_posterior_q1)
```

# ---- VIZ: Posterior of difference (p_HB - p_DO) -----------------------------#

```{r}
diff_q1_mc <- p_HB_mc - p_DO_mc

p_diff_q1 <- ggplot(data.frame(diff = diff_q1_mc[1:50000]), aes(x = diff)) +
  geom_histogram(aes(y = after_stat(density)), bins = 80,
                 fill = "#3498db", colour = "white", alpha = 0.8) +
  geom_vline(xintercept = 0,
             colour = "red", linetype = "dashed", linewidth = 1.3) +
  geom_vline(xintercept = mean(diff_q1_mc),
             colour = "#2c3e50", linewidth = 1.1) +
  annotate("label",
           x = mean(diff_q1_mc) + 0.01,
           y = max(density(diff_q1_mc[1:50000])$y) * 0.7,
           label = sprintf("Mean = %.4f\n95%% CI: [%.4f, %.4f]",
                           mean(diff_q1_mc),
                           quantile(diff_q1_mc, 0.025),
                           quantile(diff_q1_mc, 0.975)),
           hjust = 0, size = 3.8, fill = "white", label.size = 0.3) +
  labs(title    = "Q1: Posterior of (p_HB − p_DO)",
       subtitle = "Area right of 0 = P(p_HB > p_DO | data)",
       x = "Difference in hit probability (HB − DO)", y = "Density") +
  theme_minimal(base_size = 13)

print(p_diff_q1)
```

# ============================================================
# SECTION 3: Q2 — DOES MUSIC AFFECT RAILWAY VIDEO RATINGS?
#   Continuous outcome → Normal-Normal Bayesian + Welch t-test
# ============================================================

```{r}
rail_HB <- ABclean %>% filter(happyBright == 1) %>% pull(effectiveRailwayVideo)
rail_DO <- ABclean %>% filter(happyBright == 0) %>% pull(effectiveRailwayVideo)

xbar_HB <- mean(rail_HB); s_HB <- sd(rail_HB); n2_HB <- length(rail_HB)
xbar_DO <- mean(rail_DO); s_DO <- sd(rail_DO); n2_DO <- length(rail_DO)

cat(sprintf("\n=== 3. Q2 — RAILWAY VIDEO RATINGS ===\n"))
cat(sprintf("Happy & Bright : mean = %.4f, sd = %.4f, n = %d\n", xbar_HB, s_HB, n2_HB))
cat(sprintf("Dark & Ominous : mean = %.4f, sd = %.4f, n = %d\n", xbar_DO, s_DO, n2_DO))
```

# ---- 3.1 Frequentist: Welch t-test --------------------------

```{r}
freq_q2 <- t.test(rail_HB, rail_DO,
                  alternative = "two.sided", var.equal = FALSE)
cat("\n--- Frequentist Welch t-test ---\n")
print(freq_q2)
```

# ---- 3.2 Bayesian: Normal-Normal (known variance) -----------
# Prior: μ ~ N(μ_0, σ_0²)
# Choice: weakly informative — centre of scale, wide spread

```{r}
mu0         <- 5.5   # midpoint of 1–10 scale
sigma0_sq   <- 4     # prior sd = 2, diffuse over the scale
```

# 'Known' σ²: use pooled within-group variance as a plug-in estimate

```{r}
sigma_sq <- ((n2_HB - 1) * s_HB^2 + (n2_DO - 1) * s_DO^2) / (n2_HB + n2_DO - 2)
cat(sprintf("\nPooled SD used as known σ: %.4f\n", sqrt(sigma_sq)))
```

# Conjugate Normal-Normal posterior:
#   σ_n² = 1 / (1/σ_0² + n/σ²)
#   μ_n  = σ_n² × (μ_0/σ_0² + n·x̄/σ²)

```{r}
normal_normal_post <- function(n, xbar, mu0, sigma0_sq, sigma_sq) {
  var_n <- 1 / (1 / sigma0_sq + n / sigma_sq)
  mu_n  <- var_n * (mu0 / sigma0_sq + n * xbar / sigma_sq)
  list(mu = mu_n, var = var_n, sd = sqrt(var_n))
}
```

```{r}
post_HB_q2 <- normal_normal_post(n2_HB, xbar_HB, mu0, sigma0_sq, sigma_sq)
post_DO_q2 <- normal_normal_post(n2_DO, xbar_DO, mu0, sigma0_sq, sigma_sq)

cat("\n--- Bayesian Normal-Normal (full data) ---\n")
cat(sprintf("Posterior HB : N(%.4f, %.6f)  =>  95%% CI: [%.4f, %.4f]\n",
            post_HB_q2$mu, post_HB_q2$var,
            post_HB_q2$mu - 1.96 * post_HB_q2$sd,
            post_HB_q2$mu + 1.96 * post_HB_q2$sd))
cat(sprintf("Posterior DO : N(%.4f, %.6f)  =>  95%% CI: [%.4f, %.4f]\n",
            post_DO_q2$mu, post_DO_q2$var,
            post_DO_q2$mu - 1.96 * post_DO_q2$sd,
            post_DO_q2$mu + 1.96 * post_DO_q2$sd))
```

# P(μ_HB > μ_DO) via Monte Carlo

```{r}
mu_HB_mc <- rnorm(N_MC, post_HB_q2$mu, post_HB_q2$sd)
mu_DO_mc  <- rnorm(N_MC, post_DO_q2$mu, post_DO_q2$sd)
prob_HB_wins_q2 <- mean(mu_HB_mc > mu_DO_mc)
diff_q2_mc <- mu_HB_mc - mu_DO_mc

cat(sprintf("P(μ_HB > μ_DO | data)   = %.4f\n", prob_HB_wins_q2))
cat(sprintf("Posterior mean diff     = %.4f\n", mean(diff_q2_mc)))
cat(sprintf("95%% Credible Interval diff: [%.4f, %.4f]\n",
            quantile(diff_q2_mc, 0.025), quantile(diff_q2_mc, 0.975)))
```

# ---- VIZ: Posterior densities Q2 ----------------------------

```{r}
mu_grid <- seq(4.5, 8.0, length.out = 2000)

post_df_q2 <- data.frame(
  mu = rep(mu_grid, 3),
  density = c(
    dnorm(mu_grid, mu0, sqrt(sigma0_sq)),
    dnorm(mu_grid, post_HB_q2$mu, post_HB_q2$sd),
    dnorm(mu_grid, post_DO_q2$mu, post_DO_q2$sd)
  ),
  group = rep(c("Prior  N(5.5, 4)",
                "Posterior: Happy & Bright",
                "Posterior: Dark & Ominous"), each = 2000)
)
```

```{r}
p_posterior_q2 <- ggplot(post_df_q2,
                         aes(x = mu, y = density,
                             colour = group, linetype = group)) +
  geom_line(linewidth = 1.2) +
  scale_colour_manual(values = c("Prior  N(5.5, 4)"           = "#95a5a6",
                                 "Posterior: Happy & Bright"  = "#f39c12",
                                 "Posterior: Dark & Ominous"  = "#2c3e50")) +
  scale_linetype_manual(values = c("Prior  N(5.5, 4)"          = "dashed",
                                   "Posterior: Happy & Bright" = "solid",
                                   "Posterior: Dark & Ominous" = "solid")) +
  annotate("label",
           x = max(mu_grid) * 0.94,
           y = max(dnorm(mu_grid, post_HB_q2$mu, post_HB_q2$sd)) * 0.7,
           label = sprintf("P(μ_HB > μ_DO) = %.3f", prob_HB_wins_q2),
           size = 4.5, colour = "#c0392b", fill = "white", label.size = 0.4,
           fontface = "bold", hjust = 1) +
  labs(title    = "Q2: Posterior distributions — Mean railway video rating",
       subtitle = "Normal-Normal conjugate model with weakly informative prior N(5.5, 4)",
       x = "Mean effectiveness rating μ", y = "Posterior density",
       colour = "Distribution", linetype = "Distribution") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top",
        legend.text = element_text(size = 10))

print(p_posterior_q2)
```

# ---- VIZ: Posterior of difference (μ_HB - μ_DO) ------------

```{r}
p_diff_q2 <- ggplot(data.frame(diff = diff_q2_mc[1:50000]), aes(x = diff)) +
  geom_histogram(aes(y = after_stat(density)), bins = 80,
                 fill = "#9b59b6", colour = "white", alpha = 0.8) +
  geom_vline(xintercept = 0,
             colour = "red", linetype = "dashed", linewidth = 1.3) +
  geom_vline(xintercept = mean(diff_q2_mc),
             colour = "#2c3e50", linewidth = 1.1) +
  annotate("label",
           x = mean(diff_q2_mc) + 0.01,
           y = max(density(diff_q2_mc[1:50000])$y) * 0.7,
           label = sprintf("Mean = %.4f\n95%% CI: [%.4f, %.4f]",
                           mean(diff_q2_mc),
                           quantile(diff_q2_mc, 0.025),
                           quantile(diff_q2_mc, 0.975)),
           hjust = 0, size = 3.8, fill = "white", label.size = 0.3) +
  labs(title    = "Q2: Posterior of (μ_HB − μ_DO)",
       subtitle = "Area right of 0 = P(μ_HB > μ_DO | data)",
       x = "Difference in mean rating (HB − DO)", y = "Density") +
  theme_minimal(base_size = 13)

print(p_diff_q2)
```

# ============================================================
# SECTION 4: Q3 — SEQUENTIAL BAYESIAN ANALYSIS
# "Was running the A/B test for two years excessive?"
#
# IDEA: Process the 766 pairs one at a time, in chronological
# order. After each pair, update the Bayesian posterior and
# record the probability that the 'Happy & Bright' arm is
# the winner. If this probability reaches a decisive threshold
# (e.g. 0.95 or 0.05) early, the experiment could have been
# stopped sooner.
# ============================================================

```{r}
cat("\n=== 4. SEQUENTIAL BAYESIAN ANALYSIS (Running pair by pair) ===\n")

data_seq <- ABclean %>% arrange(pairwise)
n_pairs  <- max(data_seq$pairwise)
```

# Containers for sequential results

```{r}
seq_res <- data.frame(
  pair              = seq_len(n_pairs),
  prob_HB_wins_hit  = NA_real_,   # Q1: P(p_HB > p_DO)
  pmean_HB_hit      = NA_real_,   # Q1: posterior mean HB
  pmean_DO_hit      = NA_real_,   # Q1: posterior mean DO
  prob_HB_wins_rail = NA_real_,   # Q2: P(μ_HB > μ_DO)
  pmean_HB_rail     = NA_real_,   # Q2: posterior mean HB
  pmean_DO_rail     = NA_real_    # Q2: posterior mean DO
)
```

# --- Initialise priors ---
# Q1 — Beta-Binomial

```{r}
a_HB <- 1; b_HB <- 1
a_DO <- 1; b_DO <- 1
```

# Q2 — Normal-Normal (pooled σ² used throughout)

```{r}
n_HB2 <- 0; sum_HB2 <- 0
n_DO2 <- 0; sum_DO2 <- 0
```

```{r}
set.seed(42)
N_seq <- 15000  # MC draws per step (trade speed for precision)

cat("Running sequential loop over", n_pairs, "pairs...\n")

for (k in seq_len(n_pairs)) {

  pair_k <- data_seq %>% filter(pairwise == k)

  # ---- Q1 update (Beta) ----
  obs_HB_hit <- pair_k %>% filter(happyBright == 1) %>% pull(hit)
  obs_DO_hit <- pair_k %>% filter(happyBright == 0) %>% pull(hit)

  if (length(obs_HB_hit) > 0) {
    a_HB <- a_HB + sum(obs_HB_hit)
    b_HB <- b_HB + sum(1L - obs_HB_hit)
  }
  if (length(obs_DO_hit) > 0) {
    a_DO <- a_DO + sum(obs_DO_hit)
    b_DO <- b_DO + sum(1L - obs_DO_hit)
  }

  p1 <- rbeta(N_seq, a_HB, b_HB)
  p2 <- rbeta(N_seq, a_DO, b_DO)
  seq_res$prob_HB_wins_hit[k] <- mean(p1 > p2)
  seq_res$pmean_HB_hit[k]     <- a_HB / (a_HB + b_HB)
  seq_res$pmean_DO_hit[k]     <- a_DO / (a_DO + b_DO)
  
  # ---- Q2 update (Normal) ----
  
obs_HB_rail <- pair_k %>% filter(happyBright == 1) %>% pull(effectiveRailwayVideo)
  obs_DO_rail <- pair_k %>% filter(happyBright == 0) %>% pull(effectiveRailwayVideo)

  if (length(obs_HB_rail) > 0) {
    n_HB2   <- n_HB2 + length(obs_HB_rail)
    sum_HB2 <- sum_HB2 + sum(obs_HB_rail)
  }
  if (length(obs_DO_rail) > 0) {
    n_DO2   <- n_DO2 + length(obs_DO_rail)
    sum_DO2 <- sum_DO2 + sum(obs_DO_rail)
  }

  if (n_HB2 > 0 && n_DO2 > 0) {
    post_HB_s <- normal_normal_post(n_HB2, sum_HB2 / n_HB2, mu0, sigma0_sq, sigma_sq)
    post_DO_s <- normal_normal_post(n_DO2, sum_DO2 / n_DO2, mu0, sigma0_sq, sigma_sq)

    m1 <- rnorm(N_seq, post_HB_s$mu, post_HB_s$sd)
    m2 <- rnorm(N_seq, post_DO_s$mu, post_DO_s$sd)
    seq_res$prob_HB_wins_rail[k] <- mean(m1 > m2)
    seq_res$pmean_HB_rail[k]     <- post_HB_s$mu
    seq_res$pmean_DO_rail[k]     <- post_DO_s$mu
  }
}

cat("Sequential loop complete.\n")

```

# ---- When did the probability first cross decisive thresholds? ----

```{r}
THRESHOLD_HIGH <- 0.95
THRESHOLD_LOW  <- 0.05

first_cross_q1 <- seq_res %>%
  filter(prob_HB_wins_hit >= THRESHOLD_HIGH | prob_HB_wins_hit <= THRESHOLD_LOW) %>%
  slice(1)

first_cross_q2 <- seq_res %>%
  filter(!is.na(prob_HB_wins_rail) &
         (prob_HB_wins_rail >= THRESHOLD_HIGH | prob_HB_wins_rail <= THRESHOLD_LOW)) %>%
  slice(1)

cat(sprintf("\nQ1 — first pair where P crosses 0.95 or 0.05: pair %s (of 766)\n",
            ifelse(nrow(first_cross_q1) > 0, first_cross_q1$pair, "never")))
cat(sprintf("Q2 — first pair where P crosses 0.95 or 0.05: pair %s (of 766)\n",
            ifelse(nrow(first_cross_q2) > 0, first_cross_q2$pair, "never")))
```

# ---- VIZ: Sequential P(HB wins) — Q1 -----------------------

```{r}
p_seq_q1 <- ggplot(seq_res, aes(x = pair, y = prob_HB_wins_hit)) +
  geom_line(colour = "#e74c3c", linewidth = 0.7) +
  geom_hline(yintercept = THRESHOLD_HIGH,
             linetype = "dashed", colour = "#27ae60", linewidth = 0.9) +
  geom_hline(yintercept = THRESHOLD_LOW,
             linetype = "dashed", colour = "#27ae60", linewidth = 0.9) +
  geom_hline(yintercept = 0.5,
             linetype = "dotted", colour = "#7f8c8d", linewidth = 0.8) +
  {if (nrow(first_cross_q1) > 0)
    geom_vline(xintercept = first_cross_q1$pair,
               linetype = "solid", colour = "#8e44ad", linewidth = 1)} +
  annotate("text", x = 20, y = THRESHOLD_HIGH + 0.02,
           label = "0.95 — decisive evidence for HB",
           hjust = 0, size = 3.2, colour = "#27ae60") +
  annotate("text", x = 20, y = THRESHOLD_LOW - 0.02,
           label = "0.05 — decisive evidence for DO",
           hjust = 0, size = 3.2, colour = "#27ae60") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(title    = "Q1 Sequential Analysis: P(p_HB > p_DO | data so far)",
       subtitle = paste("Purple line = first pair crossing a decisive threshold;",
                        "Test ran to pair 766"),
       x = "Pairwise allocations accumulated", y = "P(Happy & Bright wins | hit)") +
  theme_minimal(base_size = 13)

print(p_seq_q1)
```

# ---- VIZ: Sequential P(HB wins) — Q2 -----------------------

```{r}
p_seq_q2 <- ggplot(seq_res %>% filter(!is.na(prob_HB_wins_rail)),
                   aes(x = pair, y = prob_HB_wins_rail)) +
  geom_line(colour = "#3498db", linewidth = 0.7) +
  geom_hline(yintercept = THRESHOLD_HIGH,
             linetype = "dashed", colour = "#27ae60", linewidth = 0.9) +
  geom_hline(yintercept = THRESHOLD_LOW,
             linetype = "dashed", colour = "#27ae60", linewidth = 0.9) +
  geom_hline(yintercept = 0.5,
             linetype = "dotted", colour = "#7f8c8d", linewidth = 0.8) +
  {if (nrow(first_cross_q2) > 0)
    geom_vline(xintercept = first_cross_q2$pair,
               linetype = "solid", colour = "#8e44ad", linewidth = 1)} +
  annotate("text", x = 20, y = THRESHOLD_HIGH + 0.02,
           label = "0.95 — decisive evidence for HB",
           hjust = 0, size = 3.2, colour = "#27ae60") +
  annotate("text", x = 20, y = THRESHOLD_LOW - 0.02,
           label = "0.05 — decisive evidence for DO",
           hjust = 0, size = 3.2, colour = "#27ae60") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(title    = "Q2 Sequential Analysis: P(μ_HB > μ_DO | data so far)",
       subtitle = paste("Purple line = first pair crossing a decisive threshold;",
                        "Test ran to pair 766"),
       x = "Pairwise allocations accumulated",
       y = "P(Happy & Bright wins | railway rating)") +
  theme_minimal(base_size = 13)

print(p_seq_q2)
```

# ---- VIZ: Evolving posterior means (Q2) ---------------------

```{r}
p_means_seq <- seq_res %>%
  filter(!is.na(pmean_HB_rail)) %>%
  pivot_longer(cols = c(pmean_HB_rail, pmean_DO_rail),
               names_to  = "group",
               values_to = "post_mean") %>%
  mutate(group = ifelse(group == "pmean_HB_rail", "Happy & Bright", "Dark & Ominous")) %>%
  ggplot(aes(x = pair, y = post_mean, colour = group)) +
  geom_line(linewidth = 0.8) +
  scale_colour_manual(values = c("Happy & Bright" = "#f39c12",
                                 "Dark & Ominous" = "#2c3e50")) +
  labs(title    = "Q2 Sequential: Evolving posterior mean (railway rating)",
       subtitle = "How the estimated group means converge as data accumulates",
       x = "Pairwise allocations accumulated",
       y = "Posterior mean effectiveness rating",
       colour = "Music group") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

print(p_means_seq)
```

# ============================================================
# FINAL SUMMARY TABLE
# ============================================================

```{r}
cat("\n\n========================================================\n")
cat("  FINAL SUMMARY\n")
cat("========================================================\n\n")

cat("[Q1] Does music affect P(children will be hit by car)?\n")
cat(sprintf("  Frequentist p-value           : %.5f\n", freq_q1$p.value))
cat(sprintf("  Interpretation                : %s\n",
            ifelse(freq_q1$p.value < 0.05,
                   "Statistically significant at 5% level",
                   "Not statistically significant at 5% level")))
cat(sprintf("  P(p_HB > p_DO | data)         : %.4f\n", prob_HB_wins_q1))
cat(sprintf("  Posterior means HB / DO       : %.4f / %.4f\n",
            a_HB_post / (a_HB_post + b_HB_post),
            a_DO_post / (a_DO_post + b_DO_post)))

cat("\n[Q2] Does music affect railway video effectiveness rating?\n")
cat(sprintf("  Frequentist p-value           : %.5f\n", freq_q2$p.value))
cat(sprintf("  Interpretation                : %s\n",
            ifelse(freq_q2$p.value < 0.05,
                   "Statistically significant at 5% level",
                   "Not statistically significant at 5% level")))
cat(sprintf("  P(μ_HB > μ_DO | data)         : %.4f\n", prob_HB_wins_q2))
cat(sprintf("  Posterior means HB / DO       : %.4f / %.4f\n",
            post_HB_q2$mu, post_DO_q2$mu))

cat("\n[Q3] Was two years excessive?\n")
cat(sprintf("  Q1 first decisive pair        : %s / 766\n",
            ifelse(nrow(first_cross_q1) > 0, first_cross_q1$pair, "never crossed")))
cat(sprintf("  Q2 first decisive pair        : %s / 766\n",
            ifelse(nrow(first_cross_q2) > 0, first_cross_q2$pair, "never crossed")))
cat("  Conclusion: see sequential plots — if a threshold was crossed\n")
cat("  well before pair 766, the trial could have been stopped earlier.\n")

cat("\n========================================================\n")
```